---
title: "WUP_04"
author: "Hailey Watkins-Oliver"
date: "fall 2024"
output:
  html_document:
    df_print: paged
---

```{r include=FALSE,echo=FALSE}
require(knitr)
require(tidyverse)
require(tigerstats)
GSScsv <- read_csv("GSScsv.csv")
```

# Introduction

<!--In this section you explain what you are trying to show.  Where did the data come from?  What is the research or other question you are trying to answer?!-->

The data is in the file GSScsv.csv and came from the 2018 General Social Survey.  The question to answer is to resolve this hypothesis.  
$H_0$ There is no difference between the average score on the WORDSUM test and level of schooling DEGREE  
$H_a$ There is a significant difference in average score on the WORDSUM test between at least two levels of DEGREE  

# Methods

<!--Mention that you intend to use chatGPT to the greatest extent possible. Give credit to chatGPT when you use it to create code in your report. Decide on your methods:  use "variable analysis" or other appropriate descriptors.  Make sure to choose at least one graphical method and at least one numerical method.!-->

# Results

<!--Divide this section into two sub-sections:  One for your descriptive  results and one for your inferential results.!-->

## Descriptive Results

### Graphical Descriptive Results

<!--Graphical results here.  Make sure to show your code.  Provide appropriate labels for axes, giving units if possible, and provide a good title for the graph, too.  Use the graphical results to describe the patterns if any that exist in the data as focused toward the research question!-->
```{r}
basicCN <- ggplot(GSScsv,aes(x=DEGREE,y=WORDSUM))

basicCN + geom_boxplot()
```

It is evident from the box plot that there is a considerable disparity between the first and third quartiles for graduates and high school students. The statement that the Sunior College first quartile is higher than the LT High School third quartile is also applicable.

### Numerical Descriptive Results

<!--Numerical results go here. Use the numerical results to describe the patterns if any that exist in the data as focused toward the research question!-->

```{r}
favstats(WORDSUM~DEGREE,data=GSScsv)
```
Graduate has a Q1 of 10, whereas LT High School's Q3 is 8.25 and OG High School's Q3 is 9, despite OG High School's large enrollment of 85 students. This is significant because the proportion of those enrolled in high school nearly equals the total number of individuals enrolled in all other degrees, indicating a significant disparity in these statistics. 


## Inferential Results

<!--State hypothesis clearly.  Make sure your discussion of the inferential test covers all the aspects that the test output produces, such as test statistic, p-value etc.  Make a decision about the null hypothesis, explain the assumptions on which the selected test/procedure was based, and why the chosen procedure satisfys the assumptions and is appropriate to answer the research question!-->

```{r}
model1 <- aov(WORDSUM~DEGREE,data=GSScsv)
summary.aov(model1)
```
The p-value is 0.000236, meaning that there is a statistical difference within these results. Further testing is required to see the given difference. 

```{r}
#we specify the margins starting at the bottom and proceeding clockwise - thus 8.7 is the left side margin to hold the names of the factors
par(mar = c(3, 15, 2, 2))
plot(TukeyHSD(model1, conf.level=0.98),las=2)
```


This graph makes this disparity clear because the values for the LT high school graduate and the high school graduate are both below 0, indicating a considerable difference between the two degrees.

# Discussion and Conclusion

<!--Discussion and conclusion here.  If you found a relationship be sure to consider whether the relationship occurs because one of the variavbles causes the other, or whether they perhasps are related for some other reason.  Watch the chapter 6 videos from the GeorgeTown videos collection.!-->

Only when we could clearly see a difference in the GRADUATE's degree from the HIGH SCHOOL DEGREE and the LT HIGH SCHOOL DEGREE did we find a relationship in the data. To properly understand this, we looked at the data, which indicated that there was a difference: the graduate's Q1 was 10, the LT High School's Q3 was 8.25, and the HIGH SCHOOL's Q3 was 9. This was demonstrated by the one-way ANOVA, which had a p-value of 0.000236 and indicated that something was occurring. We were able to identify a substantial difference between two levels in this data—the HIGH SCHOOL and LT HIGH SCHOOL—because they are both below 0. This was achieved after some tinkering with the code.

